An improved metric for the analysis of swarms

### 2017-10-20 paper.tex
Introduction and Minor text update
Intro Update and explanation of data set

%\section{Introduction\label{methods:SwarmStability}}
\emph{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists. From these studies mathematical models and algorithms have evolved. The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \emph{agents} (`robots'). The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules. These simple rules cause the agents to appear to work cooperatively. These rules can also generate movement that does not contribute to the goal of the swarm but and is a direct result of the agents attempting to move towards an optimum position within the structure. This internal movement can be identified by analysing the changes in the inter-agent interactions. This paper discusses two metrics that differ in their approach to identify these changes. The distance metric uses variations in the inter-agent spaces, as used by Navarro et al.~\cite{NIM:09}. The new metric uses the resultant magnitudes from an agents' \textit{inter-agent vectors} that are induced by an agent interacting with its neighbours.

@@

The example data extracts~(Tables \ref{tab:SampleReplusion0}, \ref{tab:SampleReplusionPositive}, \ref{tab:SampleCohesionPositive} and  \ref{tab:SampleEquilibrium}) show simulation results that are produced by the parameters listed in Table~\ref{tab:MetricPhysics1}. The simulation consists of 200 agents over a 20 second period. The cohesion of an agent pair is shown as $k_cv_c$ and the repulsion as $k_rv_r$.

@@

If the \textit{agent resultant magnitude} is negative (absolute values would prevent this analysis) the swarm's bias is to expand as the repulsion is greater than cohesion. This is seen in the chaotic stage of a swarm. If the \textit{agent resultant magnitude} is positive then the swarm is exhibiting a tendency to contract and this indicates the swarm is a cohesive entity. This could also be described as the swarm being `sticky' as the agents bias is to `pull' towards each other.

>>>>

%\section{Introduction\label{methods:SwarmStability}}
\emph{Swarming} in the animal kingdom of ants, bees, fish and birds for instance has long been studied by scientists. From these studies mathematical models and algorithms have evolved. The models and algorithms have in turn captured the interest of computing scientists who are interested in applying them to large groups of autonomous mobile \emph{agents} (`robots'). The cooperative coordination of these agents can take many forms such as following a set path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that has captured the interest of scientists is that the models and algorithms used to coordinate them are generally sets of simple rules. These simple rules cause the agents to appear to work cooperatively. These rules can also generate movement that does not contribute to the goal of the swarm this movement is a direct result of the agents attempting to move towards an optimum position within the structure. This internal movement can be identified by analysing the changes in the inter-agent interactions. Measuring this unnecessary movement allows the effectivemness of a swarming algorithm to be evaluated. If the internal movement can be reduced then stability of the swarm is improved and more of the swarms energy is focused on the goal. Also reducing the internal movement also provides a more stable platform for the deployment of sensor arrays. Both metric allow this evaluation to be caried out when all the sensor ranges are the same however the distance metric is not able to identify the internal movement accurately when the sensor ranges of each agent varies producing varyig distances even though the interactions at the magnitude level indicates the agents are in fact stable and optimum. This paper discusses two metrics and how they differ in their approach to identify these changes. The distance metric uses variations in the inter-agent spaces, as used by Navarro et al.~\cite{NIM:09}. The new metric uses the resultant magnitudes from an agents' \textit{inter-agent vectors} that are induced by an agent interacting with its neighbours which allows a more acurate analysis of a swarm when using non homogenous sensor distances The metric also identifies when a swarm is expanding or contracting.

@@

Following repeated experments in the initial deployment of a swarm it was found that using the same parameters for a swarm's environmental parameters and the cohesion and repulsion ranges that the results where always very similar. The results shown in this paper are therefore an example set typical of the results obtained. The data extracts shown in Tables \ref{tab:SampleReplusion0}, \ref{tab:SampleReplusionPositive}, \ref{tab:SampleCohesionPositive} and  \ref{tab:SampleEquilibrium}) show the simulation results that are produced by the parameters listed in Table~\ref{tab:MetricPhysics1}. The simulation consists of 200 agents over a 20 second period. The cohesion of an agent pair is shown as $k_cv_c$ and the repulsion as $k_rv_r$.

@@

If the \textit{agent resultant magnitude} is negative (absolute values would prevent this analysis) the swarm's bias is to expand as the repulsion is greater than cohesion. This is seen in the disorgansed stage of a swarm. The disorganised stage is the initial random deployment of the agents. If the \textit{agent resultant magnitude} is positive then the swarm is exhibiting a tendency to contract and this indicates the swarm is a cohesive entity. This could also be described as the swarm being `sticky' as the agents bias is to `pull' towards each other.
